`clarify`: Simulation-Based Inference for Regression Models

### 2024-05-30 clarify/Submission 2/clarify.Rmd
Reorganized for resubmission

---
title: "`clarify`: Simulation-Based Inference for Regression Models"
date: "`r Sys.Date()`"
abstract: >
  Simulation-based inference is an alternative to the delta method for computing the uncertainty around regression post-estimation (i.e., derived) quantities such as average marginal effects and adjusted predictions. It avoids several assumptions required by the delta method that are commonly violated, especially in nonlinear models, but does not require models to be refit as when using the bootstrap. It works by drawing model parameters from their joint distribution and estimating quantities of interest from each set of simulated values, which form a simulated "posterior" distribution of the quantity. `clarify` provides a simple, unified interface for performing simulation-based inference for any user-specified derived quantities as well as wrappers for common quantities of interest. `clarify` supports a large and growing number of models through its interface with the `marginaleffects` package and provides native support for multiply imputed data.
draft: true
author:  
  - name: Noah Greifer
    affiliation: Harvard University
    address:
    - Institute for Quantitative Social Science
    - Cambridge, MA
    url: ngreifer.github.io
    orcid: 0000-0003-3067-7154
    email:  ngreifer@iq.harvard.edu
  - name: Steven Worthington
    affiliation: Harvard University
    address:
    - Institute for Quantitative Social Science
    - Cambridge, MA
    email: sworthington@iq.harvard.edu
    orcid: 0000-0001-9550-5797
  - name: Stefano Iacus
    affiliation: Harvard University
    address:
    - Institute for Quantitative Social Science
    - Cambridge, MA
    email: siacus@iq.harvard.edu
    orcid: 0000-0002-4884-0047
  - name: Gary King
    affiliation: Harvard University
    address:
    - Institute for Quantitative Social Science
    - Cambridge, MA
    email: king@harvard.edu
    orcid: 0000-0002-5327-7631
type: package
output: 
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
bibliography: RJreferences.bib
---

@@

Although regression models are frequently used in empirical research to study relationships among variables, often the quantity of substantive interest is not one of the coefficients of the model, but rather a quantity derived from the coefficients, such as predicted values or average marginal effects. The usual method for estimating the uncertainty of the derived quantities is known as the "delta method", which involves two approximations: 1) that the variance of the derived quantity can be represented as a first-order Taylor series, and 2) that the estimate of the derived quantity is normally distributed[^1]. In many cases, especially with nonlinear models, these approximation can fail badly. For example, predicted probabilities close to 0 or 1 or ratios of coefficients or probabilities typically do not have normal (or even symmetrical) distributions in finite samples, and the usual Wald-type confidence interval limits produced from delta method standard errors can include values outside the domain of the quantity of interest.

[^1]: In addition, most software that uses the delta method uses numerical differentiation, which also involves an approximation.

\CRANpkg{clarify} implements an alternative to the delta method---simulation-based inference---which involves simulating a "posterior" distributions of the derived quantities. Simulation-based inference is not only often more accurate than using the delta method, it is also simpler to understand and implement, as it does not require understanding Taylor series or the calculus that underlies it. This makes it more palatable to nontechnical audiences and easier to learn for students without sacrificing statistical performance.

The methodology \CRANpkg{clarify} relies on is described in @kingMakingMostStatistical2000 and @raineyCarefulConsiderationCLARIFY2023. Simulation-based inference involves taking draws from a specified joint distribution of model parameters, computing derived quantities from these draws, and collecting the derived quantities in a "posterior" distribution, from which uncertainty measures (standard errors and confidence intervals) can be computed. This method assumes the model parameters are drawn from a multivariate Normal (or T) distribution with means at the estimated values and covariance equal to the asymptotic covariance matrix of the estimated values, a standard assumption motivated by the central limit theorem that underlies usual inference on the original model parameters. Arriving at the posterior distribution does not require taking any derivatives or making any approximations beyond those usually used for inference on model parameter estimates, except for the approximation due to Monte-Carlo error induced by sampling from a finite number of simulations. The delta method makes these same assumptions in addition to the approximations about the variance and the shape of the sampling distribution of the estimates of the derived quantities.

The nonparametric bootstrap is another alternative to the delta method for inference that does not require these approximations; it typically involves re-sampling individuals from the sample, fitting the model in each sample, and computing the quantity of interest from each model. Although bootstrapping tends to work well in practice, especially for complex and non-Normal estimators, refitting the model repeatedly can be prohibitively time-consuming and computationally expensive, especially for complicated models or large datasets. Simulation-based inference only requires the model to be fit once, and the simulations involve taking draws from a distribution produced from the single set of estimated parameters, making it much quicker in practice and allowing the user to capitalize on the already valid estimation of the model parameters.

More formally, we fit a regression model $y_i = f(x_i; \beta)$, such as a linear or other generalized linear model with model coefficients $\beta$. We assume$$
\hat{\beta} \sim \text{MVN}(\beta, \Sigma_{\hat{\beta}})
$$where $\hat{\beta}$ is the vector of estimates of $\beta$ and $\Sigma_{\hat{\beta}}$ is their asymptotic covariance matrix. We define a function $\tau(\beta)$ that represents a quantity of interest derived from the model parameters, and compute its estimate $\widehat{\tau(\beta)}$ as $\tau(\hat{\beta})$. Next, we take $M$ draws $\tilde{\beta}^{(j)}$ for $j\in(1, \dots, M)$ from a multivariate Normal distribution with mean vector $\mu = \hat{\beta}$ and covariance $\Sigma = \hat{\Sigma}_{\hat{\beta}}$, where $\hat{\Sigma}_{\hat{\beta}}$ is an estimate of the asymptotic covariance matrix of the parameter estimates. We use the distribution of $\tilde{\tau}=\tau(\tilde{\theta})$ as the "posterior" sampling distribution of $\widehat{\tau(\beta)}$, and compute its variance as$$
\sigma^2_{\widehat{\tau(\theta)}} = \frac{1}{M}\sum^M_{j=1}{(\tilde{\tau}^{(j)} - \bar{\tilde{\tau}})^2}
$$and quantile $100(1-\alpha)\%$ confidence interval limits as $\left[\tilde{\tau}_{(\frac{\alpha}{2})}, \tilde{\tau}_{(1-\frac{\alpha}{2})}\right]$ where $\tilde{\tau}_{(k)}$ is the $k$th value of $\tilde{\tau}$ when arranged in ascending order.

Similar functionality exists in the \pkg{CLARIFY} package in Stata[^2] [@tomzClarifySoftwareInterpreting2003] and used to be available in the `Zelig` R package [@imaiCommonFrameworkStatistical2008a], though there are differences in these implementations. \CRANpkg{clarify} provides additional flexibility by allowing the user to request any derived quantity, in addition to providing shortcuts for common quantities, including predictions at representative values, average marginal effects, and average dose-response functions (described below). \CRANpkg{clarify} relies on and can be seen as a companion to the \CRANpkg{marginaleffects} package, which offers similar functionality but primarily uses the delta method for calculating uncertainty (though simulation-based inference is supported in a more limited capacity as well).

@@

2.  Use `sim()` to take draws from the sampling distribution of the estimated model coefficients.

@@

```{r}
library(clarify)
```

@@

```{r}
data("lalonde", package = "MatchIt")

@@

```{r}
fit <- glm(re78_0 ~ treat * married + age + educ + race +
             nodegree + re74 + re75, data = lalonde,
           family = binomial("probit"))
```

@@

-   `dist` -- the name of the distribution from which to draw the sampled coefficients. Can be `"normal"` for a Normal distribution or `t(#)` for a t-distribution, where `#` represents the degrees of freedom. If left unspecified, `sim()` will decide on which distribution makes sense given the characteristics of the model (the decision is made by `insight::get_df()` with `type = "wald"`). Typically this does not need to be specified.

@@

`sim()` uses a random number generator to draw the sampled coefficients from the sampling distribution, so a seed should be set using `set.seed()` to ensure results are replicable across sessions.

@@

```{r}
set.seed(1234)

@@

After sampling the coefficients, we will compute derived quantities on each set of sampled coefficients and store the result, which represents the "posterior" distribution of the derived quantity, as well as on the original coefficients, which are used as the final estimates. The core functionality is provided by `sim_apply()`, which accepts a `clarify_sim` object from `sim()` and a function to compute and return one or more derived quantities, then applies that function to each set of simulated coefficients. The arguments to `sim_apply()` are below:

@@

```{r}
sim_fun1 <- function(fit) {
  predict(fit, newdata = lalonde["PSID1",], type = "response")
}
```

@@

```{r}
est1 <- sim_apply(s, FUN = sim_fun1, verbose = FALSE)

@@

```{r}
sim_fun2 <- function(coefs) {
  hispan <- unname(coefs["racehispan"])
  white <- unname(coefs["racewhite"])

@@

-   `method` -- if `ci` is `TRUE`, the method used to compute the bounds. Allowable methods include a Normal approximation (`"wald"`) or using the quantiles of the resulting distribution (`"quantile"`). The Normal approximation involves multiplying the standard deviation of the estimates (i.e., which functions like the standard error of the sampling distribution) by the critical Z-statistic computed using `(1-level)/2` to create a symmetric margin of error around the point estimate. The default is `"quantile"` to instead use quantile-based bounds, which are more appropriate when the distribution is non-Normal. However, quantile-based bounds may require more simulations to stabilize.

-   `reference` -- whether to display a normal density over the plot for each estimate. The default is `FALSE` to omit it.

@@

```{r, fig.width=4}
plot(est1, reference = TRUE, ci = FALSE)
```

Overlaid on the plot in red is a Normal distribution with the same mean and standard deviation as the simulated values; this is requested by setting `reference = TRUE`. From the plot, one can see that the distribution of simulated values is non-Normal, asymmetrical, and not centered around the estimate, with no values falling above 1 because the outcome is a predicted probability. Given its non-Normality, the quantile-based bounds are clearly more appropriate than those resulting from the Normal approximation, as the bounds computed from the Normal approximation would be outside the bounds of the estimate. The plot itself is a `ggplot` object that can be modified using `ggplot2` syntax.

@@

-   `null` -- an optional argument specifying the null value in a hypothesis test for the estimates. If specified, a p-value will be computed using either a standard Z-test (if `method` is `"wald"`) or an inversion of the uncertainty interval (described below). The default is not to display any p-values.

Inverting the uncertainty interval involves finding the smallest confidence level such that the null value is within the confidence bounds. The p-value for the test is one minus this level. It is only valid as a p-value when the simulated distribution of the estimates differs from its true sampling distribution under the null value by a location shift (which can be violated when the distribution of the estimates is asymmetric). However, because the p-values are invariant to monotonic transformations of the estimates and null value, as long as the distribution of *some* monotonic transformation of the estimate is a location shift from its sampling distribution under the null hypothesis, the p-values will be valid.

@@

```{r}
summary(est1)
```

A benefit of quantile-based confidence bounds is that the bounds will always be in the domain of the quantity of interest, e.g., bounds for probabilities will not be above 1 or below 0 if the model always generates predicted probabilities between 0 and 1. In contrast, Wald-type symmetric confidence intervals (including those produced by the delta method) can include values outside the domain of the quantity of interest. While it is possible to compute Wald-type confidence interval bounds that respect the domain boundaries, e.g., by computing the confidence interval around the logit of the predicted probability and using the inverse logit of the bounds as the bound on the probability, quantile confidence intervals do not require the user to know which transformation of the quantity of interest will yield appropriate bounds.

@@

```{r, fig.width=4}
plot(est2, reference = TRUE, ci = FALSE)

@@

The uncertainty intervals and p-values in the `summary()` output are computed using the Normal approximation because we set `method = "wald"`, and the p-value for the test that our estimate is equal to 0 is returned because we set `null = 0`. Note that the Normal approximation should be used only when the simulated posterior distribution is both close to Normal and centered around the estimate (i.e., when the mean of the simulated values [red vertical line] coincides with the estimate computed on the original coefficients [black vertical line]).

@@

```{r}
est3 <- sim_setx(s,
                 x = list(treat = 0:1,
                          re75 = c(0, 20000),
                          race = "black"),
                 verbose = FALSE)
```

@@

```{r}
summary(est3)
```

@@

```{r}
attr(est3, "setx")
```

@@

```{r, fig.width=5}
plot(est3, var = "re75", ci = FALSE)
```

One can see again how a delta method or Normal approximation may not have yielded valid uncertainty intervals given the non-Normality of the distributions.

@@

```{r}
est4 <- sim_setx(s,
                 x = list(treat = 0:1,
                          re75 = seq(0, 20000, by = 2000),
                          race = "black"),
                 verbose = FALSE)
```

@@

```{r, fig.width=5}
plot(est4)
```

@@

```{r}
est5 <- sim_setx(s,
                 x = list(treat = 0, re75 = 0),
                 x1 = list(treat = 1, re75 = 0),
                 verbose = FALSE)
```

@@

```{r}
summary(est5)
```

@@

Using predicted values and effects at representative values is one way to summarize regression models, but another way is to compute average adjusted predictions (AAPs), contrasts of AAPs, and average marginal effects (AMEs). The definitions for these terms may vary and the names for these concepts differ across sources, but here we define AAPs as the average of the predicted values for all units after setting one predictor to a chosen value, and we define AMEs for binary predictors as the contrast of two AAPs and for continuous predictors as the average of instantaneous rate of change in the AAP corresponding to a small change in the predictor from its observed values across all units[^3] [@longRegressionModelsCategorical2014].

[^3]: In \CRANpkg{marginaleffects}, AAPs are computed using `avg_predictions()`, AMEs for binary variables are computed using `avg_comparisons()`, and AMEs for continuous variables are computed using `avg_slopes()`. AAPs are sometimes known as average "counterfactual" predictions.

@@

```{r}
est6 <- sim_ame(s,
                var = "treat",
                subset = treat == 1,
                contrast = "rr",
                verbose = FALSE)
```

@@

```{r}
summary(est6, null = c(`RR` = 1))
```

@@

```{r}
est7 <- sim_ame(s,
                var = "age",
                verbose = FALSE)
```

@@

```{r}
summary(est7)
```

@@

```{r}
est6b <- sim_ame(s,
                 var = "treat",
                 subset = treat == 1,
                 by = ~married,
                 contrast = "rr",
                 verbose = FALSE)

@@

```{r}
age_seq <- seq(18, 50, by = 2)

@@

```{r, fig.width=5}
plot(est8)
```

@@

```{r}
summary(est8, parm = 1:4)
```

@@

```{r}
est9 <- sim_adrf(s,
                 var = "age",
                 contrast = "amef",
                 at = age_seq,
                 verbose = FALSE)
```

@@

```{r, fig.width=5}
plot(est9)
```

@@

`transform()` is a generic function in R that is typically used to create a new variable in a data frame that is a function of other columns. For example, to compute the binary outcome we used in our model, we could have run the following[^4]:

[^4]: Users familiar with the `tidyverse` will note the similarities between `transform()` and `dplyr::mutate()`; only `transform()` can be used with `clarify_est` objects.

```{r}
lalonde <- transform(lalonde,
                     re78_0 = ifelse(re78 == 0, 1, 0))
```

@@

```{r}
est6 <- transform(est6,
                  RD = `E[Y(1)]` - `E[Y(0)]`)
```

Note that we used tics (\`) around the names of the AAPs; this is necessary when they contain special characters like parentheses or brackets.

@@

```{r}
summary(est6, null = c(`RR` = 1, `RD` = 0))
```

One benefit of using simulation-based inference with p-values computed from inverting the confidence intervals is that the p-values for the risk difference and risk ratio (and any other effect measure for comparing a pair of values) will always exactly align, thereby ensuring inference does not depend on the effect measure used. In contrast, Wald-type inference (based either on the simulation-derived standard error or the delta method) is not invariant to transformations of the quantity of interest.

@@

```{r}
est10 <- cbind(est10b, est10h)
summary(est10)
```

>>>>

---
title: "`clarify`: Simulation-Based Inference for Regression Models"
date: "`r Sys.Date()`"
abstract: >
  Simulation-based inference is an alternative to the delta method for computing the uncertainty around regression post-estimation (i.e., derived) quantities such as average marginal effects, average adjusted predictions, and other functions of model parameters. It works by drawing model parameters from their joint distribution and estimating quantities of interest from each set of simulated values, which form a simulated "posterior" distribution of the quantity from which confidence intervals can be computed. `clarify` provides a simple, unified interface for performing simulation-based inference for any user-specified derived quantities as well as wrappers for common quantities of interest. `clarify` supports a large and growing number of models through its interface with the `marginaleffects` package and provides native support for multiply imputed data.
draft: true
author:  
  - name: Noah Greifer
    affiliation: Harvard University
    address:
    - Institute for Quantitative Social Science
    - Cambridge, MA
    url: ngreifer.github.io
    orcid: 0000-0003-3067-7154
    email:  ngreifer@iq.harvard.edu
  - name: Steven Worthington
    affiliation: Harvard University
    address:
    - Institute for Quantitative Social Science
    - Cambridge, MA
    email: sworthington@iq.harvard.edu
    orcid: 0000-0001-9550-5797
  - name: Stefano Iacus
    affiliation: Harvard University
    address:
    - Institute for Quantitative Social Science
    - Cambridge, MA
    email: siacus@iq.harvard.edu
    orcid: 0000-0002-4884-0047
  - name: Gary King
    affiliation: Harvard University
    address:
    - Institute for Quantitative Social Science
    - Cambridge, MA
    email: king@harvard.edu
    orcid: 0000-0002-5327-7631
type: package
output: 
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
bibliography: RJreferences.bib
---

@@

Although regression models are frequently used in empirical research to study relationships among variables, often the quantity of substantive interest is not one of the coefficients of the model, but rather a quantity derived from the coefficients, such as predicted values or average marginal effects. Quantifying the uncertainty of these derived quantities (i.e., computing standard errors, confidence intervals, and p-values) requires additional processing. Several methods of doing so exist, including the delta method, the bootstrap, and simulation-based inference. \CRANpkg{clarify} implements simulation-based inference, which we describe below along with these other methods.

The delta method involves computing a first-order Taylor series approximation to the variance of the derived quantity, and standard Wald-based inference relies on computing quantiles based on the Normal distribution and using them to compute p-values and confidence intervals. \CRANpkg{clarify} implements an alternative to the delta method---simulation-based inference---which involves simulating a "posterior" distribution of the derived quantities. Simulation-based inference does not require understanding Taylor series or the calculus that underlies it, which can make it more palatable to nontechnical audiences and easier to learn for students without necessarily sacrificing statistical performance [@kingMakingMostStatistical2000; @zelnerUsingSimulationInterpret2009]. Some studies have found that simulation-based inference performs as well or better than the delta method for computing derived quantities (i.e., with respect to achieving close to nominal coverage for confidence intervals), especially for complicated derived quantities and in smaller samples [@mackinnonConfidenceLimitsIndirect2004; @holeComparisonApproachesEstimating2007; @herronPostestimationUncertaintyLimited1999]. Its empirical performance has been particularly well studied in the context of mediation analysis, in which the quantities of interest are products and ratios of regression coefficients, where it has been shown to perform well relative to the delta method due to the non-Normality of these quantities [@tofighiMonteCarloConfidence2016; @preacherAdvantagesMonteCarlo2012].

The methodology \CRANpkg{clarify} relies on was developed by @krinsky1986 and is described in @kingMakingMostStatistical2000 and @herronPostestimationUncertaintyLimited1999. Simulation-based inference involves taking draws from a specified joint distribution of model parameters, computing derived quantities from these draws, and collecting the derived quantities in a "posterior" distribution, from which uncertainty measures (standard errors and confidence intervals) can be computed. This method assumes the model parameters are drawn from a multivariate Normal (or T) distribution with means at the estimated values and covariance equal to the asymptotic covariance matrix of the estimated values, a standard assumption motivated by the central limit theorem that underlies usual inference on the original model parameters. Arriving at the posterior distribution does not require taking any derivatives or making any approximations beyond those usually used for inference on model parameter estimates, except for the approximation due to Monte Carlo error induced by sampling from a finite number of simulations (which can always be reduced by increasing the number of draws at the cost of increased computing time).

The nonparametric bootstrap is another alternative to the delta method for inference that does not require its analytic approximations [@efronBootstrapMethodsStandard1986]; bootstrapping typically involves re-sampling individuals from the sample, fitting the model in each bootstrap sample, and computing the quantity of interest from each model. Although bootstrapping tends to work well in practice, especially for complex and non-Normal estimators, refitting the model repeatedly can be prohibitively time-consuming and computationally expensive, especially for complicated models or large datasets. Simulation-based inference only requires the model to be fit once, and the simulations involve taking draws from a distribution produced from the single set of estimated parameters, making it much quicker in practice and allowing the user to capitalize on the already valid estimation of the model parameters. Methods for computing valid confidence intervals in cases when the quantity of interest has a complicated distribution are better developed when bootstrapping, however [@efronBootstrapMethodsStandard1986].

More formally, we fit a regression model $y_i = f(x_i; \beta)$, such as a linear or other generalized linear model with model coefficients $\beta$. We assume$$
\hat{\beta} \sim \text{MVN}(\beta, \Sigma_{\hat{\beta}})
$$where $\hat{\beta}$ is the vector of estimates of $\beta$ and $\Sigma_{\hat{\beta}}$ is their asymptotic covariance matrix. We define a function $\tau(\beta)$ that represents a quantity of interest derived from the model parameters, and compute its estimate $\widehat{\tau(\beta)}$ as $\tau(\hat{\beta})$. To perform simulation-based inference, we take $M$ draws $\tilde{\beta}^{(j)}$ for $j\in(1, \dots, M)$ from a multivariate Normal distribution with mean vector $\mu = \hat{\beta}$ and covariance $\Sigma = \hat{\Sigma}_{\hat{\beta}}$, where $\hat{\Sigma}_{\hat{\beta}}$ is an estimate of the asymptotic covariance matrix of the parameter estimates. We use the distribution of $\tilde{\tau}=\tau(\tilde{\beta})$ as the "posterior" distribution of $\widehat{\tau(\beta)}$, and compute its variance as$$
\hat\sigma^2_{\widehat{\tau(\beta)}} = \frac{1}{M-1}\sum^M_{j=1}{(\tilde{\tau}^{(j)} - \bar{\tilde{\tau}})^2}
$$and quantile $100(1-\alpha)\%$ confidence interval limits as $\left[\tilde{\tau}_{(\frac{\alpha}{2})}, \tilde{\tau}_{(1-\frac{\alpha}{2})}\right]$ where $\tilde{\tau}_{(q)}$ is the $q$th value of $\tilde{\tau}$ when arranged in ascending order (i.e., the $q$th quantile of the empirical cumulative distribution function). Simulation-based Wald-type confidence intervals can be computed as $\left[\tau(\hat{\beta}) + \hat\sigma_{\widehat{\tau(\beta)}} Z_{\frac{\alpha}{2}}, \tau(\hat{\beta}) + \hat\sigma_{\widehat{\tau(\beta)}} Z_{1-\frac{\alpha}{2}} \right]$, where $Z_q$ is the $q$th quantile of a standard Normal distribution. The delta method-based Wald-type confidence intervals use this formula but with the first-order Taylor approximation to the asymptotic variance: $\hat\sigma^2_{\widehat{\tau(\beta)}}=\nabla\tau(\hat{\beta}) \Sigma_{\hat{\beta}} \nabla\tau'(\hat{\beta})$, where $\nabla\tau(\hat{\beta})$ is the gradient of $\tau(\beta)$ with respect to $\beta$ evaluated at $\hat{\beta}$.

To compute a p-value for a hypothesis test involving the quantity of interest, i.e., $\text{H}_0: \tau (\beta) = \tau_0$ with a given null value $\tau_0$, we can invert the confidence interval [@thulinModernStatisticsWrangling2021]; that is, we find the largest value of $\alpha$ such that $\tau_0$ is within the confidence interval and use that $\alpha$ as the p-value for the test. For Wald-based inference (either using the simulation-based variance or delta method-based variance), this is equivalent to performing a standard two-sided Z-test using the test statistic $Z=\left(\tau(\hat\beta)-\tau_0\right) / \hat\sigma_{\widehat{\tau(\beta)}}$. One benefit of using the quantile p-values for inference is that equivalent tests of the same hypothesis will always yield identical p-values; for example, testing the equality of two derived quantities will yield the same p-value when comparing the difference between the quantities against a null hypothesis of 0 and the ratio of the quantities against a null hypothesis of 1, as each of these hypothesis is true if and only if the other is true.

One would expect simulation-based quantile inference, simulation-based Wald inference, and delta method-based Wald inference to align when the posterior is Normally distributed around the estimate, in which case any discrepancies would be due to Monte Carlo error in the simulated values (and therefore would shrink with increasing draws). However, for low values of $\alpha$, it may require many draws for the simulation-based intervals to stabilize; delta method-based intervals are not subject to this error. There are a few cases in which the results might diverge: in some cases, the first-order Taylor series approximation to the variance may be poor, though in practice the approximation error is small and shrinks quickly with increasing sample size. When the posterior distribution is non-Normal but symmetric around the estimate, the quantile intervals may be more accurate (i.e., in the sense of achieving closer to nominal coverage) because they do not rely on quantiles from the Normal distribution [@tofighiMonteCarloConfidence2016].

Another potential advantage quantile intervals can have over Wald-intervals is that when *some* monotonic transformation of the estimate has a symmetric distribution centered around the transformed estimate, the quantile intervals can achieve correct coverage without requiring knowledge of *which* transformation is required [@efronBootstrapMethodsStandard1986]; this is true of the quantile-based p-values as well. When the distribution is not centered around the estimate and no monotonic transformation will make it so, though, neither quantile-based nor Wald-based intervals would be expected to perform well, and quantile intervals could yield even worse coverage than Wald-based intervals, a phenomenon that occurs in the context of bootstrapping [@efronBootstrapMethodsStandard1986][^1]. An informal falsification test for whether such a monotonic transformation exists is whether the median of the simulated estimates is aligned with the point estimate; if it is not, there is no monotonic transformation that will yield a symmetric quantile interval with the desired coverage.

[^1]: We thank an anonymous reviewer for pointing out a scenario in which this could occur: for a quantity of interest with a right-skewed sampling distribution, one would prefer an estimate to the right of the quantity's true value to have a confidence interval skewed to left to capture the bulk of the sampling distribution, but in practice a quantile confidence interval would also be skewed to the right. While a symmetric Wald-based interval may not have adequate coverage, the quantile-based interval could perform even worse.

## Related Software

Similar functionality exists in the \pkg{CLARIFY} package in Stata[^2] [@tomzClarifySoftwareInterpreting2003] and used to be available in the `Zelig` R package [@imaiCommonFrameworkStatistical2008a], though there are differences in these implementations. \CRANpkg{clarify} provides additional flexibility by allowing the user to request any derived quantity, in addition to providing shortcuts for common quantities, including predictions at representative values, average marginal effects, and average dose-response functions (described below). \CRANpkg{clarify} relies on and can be seen as a companion to the \CRANpkg{marginaleffects} package [@arel-bundockMarginaleffectsPredictionsComparisons2024], which offers similar functionality but primarily uses the delta method for calculating uncertainty (though simulation-based inference is supported in a more limited capacity as well).

@@

2.  Use `sim()` to take draws from the multivariate distribution of the estimated model coefficients.

@@

```{r library}
library(clarify)
```

@@

```{r data}
data("lalonde", package = "MatchIt")

@@

```{r fit}
fit <- glm(re78_0 ~ treat * married + age + educ + race +
             nodegree + re74 + re75, data = lalonde,
           family = binomial("probit"))
```

@@

-   `dist` -- the name of the distribution from which to draw the sampled coefficients. Can be `"normal"` for a Normal distribution or `t(#)` for a T-distribution, where `#` represents the degrees of freedom. If left unspecified, `sim()` will decide on which distribution makes sense given the characteristics of the model (the decision is made by `insight::get_df()` with `type = "wald"`). Typically this does not need to be specified.

@@

`sim()` uses a random number generator to draw the sampled coefficients from the sampling distribution, so a seed should be set using `set.seed()` to ensure results are replicable across sessions. Using more iterations (i.e., increasing `n`) yields results that will be more stable across runs even when a seed is not set.

@@

```{r sim}
set.seed(1234)

@@

After sampling the coefficients, one can compute derived quantities on each set of sampled coefficients and store the result, which represents the "posterior" distribution of the derived quantity, as well as on the original coefficients, which are used as the final estimates. The core functionality is provided by `sim_apply()`, which accepts a `clarify_sim` object from `sim()` and a function to compute and return one or more derived quantities, then applies that function to each set of simulated coefficients. The arguments to `sim_apply()` are below:

@@

```{r sim_fun1}
sim_fun1 <- function(fit) {
  predict(fit, newdata = lalonde["PSID1",], type = "response")
}
```

@@

```{r est1}
est1 <- sim_apply(s, FUN = sim_fun1, verbose = FALSE)

@@

```{r sim_fun2}
sim_fun2 <- function(coefs) {
  hispan <- unname(coefs["racehispan"])
  white <- unname(coefs["racewhite"])

@@

-   `method` -- if `ci` is `TRUE`, the method used to compute the bounds. Allowable methods include thea Normal approximation (`"wald"`) or using the quantiles of the resulting distribution (`"quantile"`). The Normal approximation involves multiplying the standard deviation of the estimates (i.e., which functions like the standard error of the sampling distribution) by the critical Z-statistic computed using `(1-level)/2` to create a symmetric margin of error around the point estimate. The default is `"quantile"` to instead use quantile-based bounds.

-   `reference` -- whether to display a normal density over the plot for each estimate and an indicator line for the median of the estimate. The default is `FALSE` to omit them.

@@

```{r plot1, fig.width=4}
plot(est1, reference = TRUE, ci = FALSE)
```

Overlaid on the plot in red is a Normal distribution with the same mean and standard deviation as the simulated values; this is requested by setting `reference = TRUE`. From the plot, one can see that the distribution of simulated values is non-Normal, asymmetrical, and not centered around the estimate, with no values falling above 1 because the outcome is a predicted probability. Given its non-Normality, the quantile-based bounds may be more appropriate than those resulting from the Normal approximation, as the bounds computed from the Normal approximation would be outside the bounds of the estimate. The blue reference line for the median of the estimates is close to the point estimate, suggesting it is possible for a monotonic transformation to have a symmetric distribution around the estimate[^3]. The plot itself is a `ggplot` object that can be modified using `ggplot2` syntax.

[^3]: In fact, we know the inverse link function for the model (i.e., the Normal distribution function `qnorm()`) is such a transformation in this case; \CRANpkg{marginaleffects} and other packages that implement the delta method for confidence intervals around model predictions typically automatically compute confidence intervals on the transformed predictions before transforming the intervals back using the model's link function. As long as such a transformation exists, the quantile intervals will be as valid as those that rely on transforming and back-transforming.

@@

-   `null` -- an optional argument specifying the desired null value in a hypothesis test for the estimates. If specified, a p-value will be computed using either a standard Z-test (if `method` is `"wald"`) or an inversion of the uncertainty interval (described below). The default is not to display any p-values.

@@

```{r summary1}
summary(est1)
```

@@

```{r plot2, fig.width=4}
plot(est2, reference = TRUE, ci = FALSE)

@@

The uncertainty intervals and p-values in the `summary()` output are computed using the Normal approximation because we set `method = "wald"`, and the p-value for the test that our estimate is equal to 0 is returned because we set `null = 0`. Note that the Normal approximation should be used only when the simulated posterior distribution is both close to Normal and centered around the estimate (i.e., when the mean of the simulated values [red vertical line] coincides with the estimate computed on the original coefficients [black vertical line]). In such cases, however, the delta method will likely perform as well, if not better, and all of its other benefits apply (i.e., it is computationally quicker and not subject to Monte Carlo error).

@@

```{r setx3}
est3 <- sim_setx(s,
                 x = list(treat = 0:1,
                          re75 = c(0, 20000),
                          race = "black"),
                 verbose = FALSE)
```

@@

```{r summary3}
summary(est3)
```

@@

```{r attr3}
attr(est3, "setx")
```

@@

```{r plot3, fig.width=5}
plot(est3, var = "re75", ci = FALSE)
```

One can see again how a delta method or Normal approximation may have yielded uncertainty intervals outside the plausible range of the estimate without applying knowledge of the correct transformation to use to avoid doing so.

@@

```{r setx4}
est4 <- sim_setx(s,
                 x = list(treat = 0:1,
                          re75 = seq(0, 20000, by = 2000),
                          race = "black"),
                 verbose = FALSE)
```

@@

```{r plot4, fig.width=5}
plot(est4)
```

@@

```{r setx5}
est5 <- sim_setx(s,
                 x = list(treat = 0, re75 = 0),
                 x1 = list(treat = 1, re75 = 0),
                 verbose = FALSE)
```

@@

```{r summary5}
summary(est5)
```

@@

Using predicted values and effects at representative values is one way to summarize regression models, but another way is to compute average adjusted predictions (AAPs), contrasts of AAPs, and average marginal effects (AMEs). The definitions for these terms may vary and the names for these concepts differ across sources, but here we define AAPs as the average of the predicted values for all units after setting one predictor to a chosen value, and we define AMEs for binary predictors as the contrast of two AAPs and for continuous predictors as the average of instantaneous rate of change in the AAP corresponding to a small change in the predictor from its observed values across all units[^4] [@longRegressionModelsCategorical2014].

[^4]: In \CRANpkg{marginaleffects}, AAPs are computed using `avg_predictions()`, AMEs for binary variables are computed using `avg_comparisons()`, and AMEs for continuous variables are computed using `avg_slopes()`. AAPs are sometimes known as average "counterfactual" predictions.

@@

```{r ame6}
est6 <- sim_ame(s,
                var = "treat",
                subset = treat == 1,
                contrast = "rr",
                verbose = FALSE)
```

@@

```{r summary6}
summary(est6, null = c(`RR` = 1))
```

@@

```{r ame7}
est7 <- sim_ame(s,
                var = "age",
                verbose = FALSE)
```

@@

```{r summary7}
summary(est7)
```

@@

```{r ame6b}
est6b <- sim_ame(s,
                 var = "treat",
                 subset = treat == 1,
                 by = ~married,
                 contrast = "rr",
                 verbose = FALSE)

@@

```{r adrf8}
age_seq <- seq(18, 50, by = 2)

@@

```{r plot8, fig.width=5}
plot(est8)
```

@@

```{r summary8}
summary(est8, parm = 1:4)
```

@@

```{r adrf9}
est9 <- sim_adrf(s,
                 var = "age",
                 contrast = "amef",
                 at = age_seq,
                 verbose = FALSE)
```

@@

```{r plot9, fig.width=5}
plot(est9)
```

@@

`transform()` is a generic function in R that is typically used to create a new variable in a data frame that is a function of other columns. For example, to compute the binary outcome we used in our model, we could have run the following[^5]:

[^5]: Users familiar with the `tidyverse` will note the similarities between `transform()` and `dplyr::mutate()`; only `transform()` can be used with `clarify_est` objects.

```{r transform_demo}
lalonde <- transform(lalonde,
                     re78_0 = ifelse(re78 == 0, 1, 0))
```

@@

```{r transform6}
est6 <- transform(est6,
                  RD = `E[Y(1)]` - `E[Y(0)]`)
```

Note that we used tics (`` ` ``) around the names of the AAPs; this is necessary when they contain special characters like parentheses or brackets. An alternative is to use the shortcut names `.b#`, where `#` is replaced with a number (e.g., as `.b1`, `.b2`, etc.) corresponding to the index of the quantity referenced. For example, because `E[Y(1)]`and`E[Y(0)]` are the second and first computed quantities, respectively, the above code could be replaced with

```{r transform6b, eval = FALSE}
est6 <- transform(est6,
                  RD = .b2 - .b1)
```

which will yield identical results[^6].

[^6]: Note that if a quantity is named `.b#`, e.g., `.b1`, it can only be referred to using the positional shortcut and not its named. That is, the positional shortcut takes precedence over the names of the quantities.

@@

```{r summary6_transform}
summary(est6, null = c(`RR` = 1, `RD` = 0))
```

As mentioned previously, one benefit of using simulation-based inference with p-values computed from inverting the confidence intervals is that the p-values for testing the same hypothesis with the risk difference and risk ratio (and any other effect measure for comparing a pair of values) will always exactly align, thereby ensuring inference does not depend on the effect measure used. In contrast, Wald-type inference (based on either the simulation-derived or delta method standard error) is not invariant to transformations of the quantity of interest.

@@

```{r cbind10}
est10 <- cbind(est10b, est10h)
summary(est10)
```
